Dependent Object Types

### 2012-09-07 doc/candidacy/slides.tex
proposal slide

\begin{frame}[fragile]{${\nu}Obj$: Example (encoding of monomorphic functions)}
A $\lambda$-abstraction $\lambda ( x : T )\; t$ is represented as a
class with an abstract member \code!arg! for the function argument:
\begin{lstlisting}
[x: {arg: T}| fun = [res = $t'$]]
\end{lstlisting}
$t'$ corresponds to term $t$ in which all occurrences of $x$ are
replaced by \code!$x$.arg!. We cannot directly access \code!arg! on
the right-hand side of \code!fun!: this wouldn't be ``contractive'',
because \code!arg! might not be initialized before \code!fun!. Hence,
we pack $t'$ in another class to ensure contractiveness. Now, an
application $g(e)$ gets decomposed into three subsequent steps:
\begin{lstlisting}
$\nu\,$g$_{app}$  <- g & [arg = e];
$\nu\,$g$_{eval}$ <- g$_{app}$.fun;
g$_{eval}$.res
\end{lstlisting}
First we instantiate function \code!g! with a concrete argument
yielding a thunk \code!g$_{app}$!. Then we evaluate this thunk by
creating an instance \code!g$_{eval}$! of it. Finally we extract the
result by querying field \code!res! of \code!g$_{eval}$!.
\end{frame}

\begin{frame}[fragile]{${\nu}Obj$: Discussion}
\begin{itemize}
\item conflate the concepts of compound types (which inherit the
  members of several parent types) and mixin composition (which build
  classes from other classes and traits)
\item mixin composition is not commutative, unlike classical
  intersection types
\item in Scala, least upper bounds and greatest lower bounds do not always exist, e.g:
\end{itemize}
\begin{minted}{scala}
  trait A { type T <: A }
  trait B { type T <: B }
  // glb is approximated by an infinite sequence
  A with B { type T <: A with B { type T <: A with B {
    type T <: ...
  }}}
\end{minted}
\end{frame}

\section{Proposal}

>>>>

% can you make this slide less cluttered?
\begin{frame}[fragile]{${\nu}Obj$: Example (encoding of monomorphic functions)}
A $\lambda$-abstraction $\lambda ( x : T )\; t$ is represented as a
class with an abstract member \code!arg! for the function argument:
\begin{lstlisting}
[x: {arg: T}| fun = [res = $t'$]]
\end{lstlisting}
$t'$ corresponds to term $t$ in which all occurrences of $x$ are
replaced by \code!$x$.arg!. We cannot directly access \code!arg! on
the right-hand side of \code!fun!: this wouldn't be ``contractive'',
because \code!arg! might not be initialized before \code!fun!. Hence,
we pack $t'$ in another class to ensure contractiveness. Now, an
application $g(e)$ gets decomposed into three subsequent steps:
\begin{lstlisting}
$\nu\,$g$_{app}$  <- g & [arg = e];
$\nu\,$g$_{eval}$ <- g$_{app}$.fun;
g$_{eval}$.res
\end{lstlisting}
First we instantiate function \code!g! with a concrete argument
yielding a thunk \code!g$_{app}$!. Then we evaluate this thunk by
creating an instance \code!g$_{eval}$! of it. Finally we extract the
result by querying field \code!res! of \code!g$_{eval}$!.
\end{frame}

\begin{frame}[fragile]{${\nu}Obj$: Discussion}
\begin{itemize}
\item conflate the concepts of compound types (which inherit the
  members of several parent types) and mixin composition (which build
  classes from other classes and traits)
\item mixin composition is not commutative, unlike classical
  intersection types
\item in Scala, least upper bounds and greatest lower bounds do not always exist, e.g: %can you translate this to nuObj?
\end{itemize}
\begin{minted}{scala}
  trait A { type T <: A }
  trait B { type T <: B }
  // glb is approximated by an infinite sequence
  A with B { type T <: A with B { type T <: A with B {
    type T <: ...
  }}}
\end{minted}
\end{frame}

\section{Proposal}
\begin{frame}{DOT: Dependent Object Types}
\begin{itemize}
\item core calculus for modeling {\it path-dependent types}
\item we've seen path-dependent types arise in three settings:
\begin{itemize}
\item ML-like module systems
\item virtual classes
\item ${\nu}Obj$ / Scala
\end{itemize}
\item DOT aims to bring more uniformity and simplicity to Scala
\begin{itemize}
\item replace Scala's compound types with classical intersection types
\item complement calculus with classical union types
\item intersections and unions form a lattice wrt subtyping
\end{itemize}
\end{itemize}
\end{frame}
